---
title: "Using RMarkdown"
author: ''
date: "July 12, 2014"
output:
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r include=FALSE}
# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

### Text

Text can decorated with **bold** or *italics*.  It is also possible to 

* create [links](http://rmarkdown.rstudio.com/) 
* include mathematics like $e=mc^2$ or 
$$y = \beta_0 + \beta_1 x_1 + \beta_2 x_2$$
* embed images 

Be sure to put a space after the * when you are creating bullets and a space after # when 
creating section headers, but **not** between $ and the mathematical formulas.

### Graphics

If the code of an R chunk produces a plot, this plot can be displayed in the resulting file.

```{r}
plot(mtcars$mpg ~ mtcars$hp)
```

### R output
Other forms of R output are also displayed as they are produced.

```{r}
summary(mtcars$mpg)
```

You can also use "chunk options" to modify how your R chunks work.  We can hide the code:

```{r, echo = FALSE}
summary(mtcars$mpg)
```

Or we can hide the output:

```{r, eval=FALSE}
summary(mtcars$mpg)
```

### In-text Code

In a similar vein, you can refer to R code in a paragraph by setting it off with single
ticks, say if we wanted to mentioned that we had used the `summary()` function.  You can
also put R output directly into text, as in: the mean number of births was 
`r mean(mtcars$mpg)`.

### Destination formats

This file can be knit to HTML, PDF, or Word.  In RStudio, just select the desired 
output file type and click on `Knit HTML`, `Knit PDF`, or `Knit Word`.  Use the 
dropdown menu next to that to change the desired file type.

* * *

### On your own
Please create a new markdown file called "Brief Bio" with two sections.

#### 1. Hometown
Provide your hometown and include a photo of it.  Also print
out some basic demographic data on your home state by using the following code.

```{r, eval=FALSE}
data(states)
states.x77[rowNumber,]
```

You'll need to replace `rowNumber` with the row number of your state.  Also, please
hide the R code and provide only the output.

#### 2. Food
Create a numbered list of your top three favorite restaurants of all time and provide links
to their [Yelp](http://www.yelp.com/) page.
