---
title: "Writing Functions"
output:
  ioslides_presentation:
    incremental: true
---


## Recall {.build}

*"Everything that happens is a function call.  Everything that exists is an object."*

## Including functions! {.build}

```{r}
g1 <- function(x) {
  return(x + 1)
}
g1(5)
```


## The output {.build}

Use `return()` to be explicit, otherwise R will return the last value computed.

```{r}
g1 <- function(x) {
  return(x + 1)
}
g2 <- function(x) {
  x + 1
}
g1(5)
g2(5)
```


## The Input {.build}

You can list as many arguments as desired, specifying default values with `=`
and flexible additional arguments with an ellipsis `...`

```{r}
cylVol <- function(h = 1, r = 1) {
  pi * r^2 * h
}
cylVol(3, 4)
cylVol(3)
```


## Scoping {.build}

Functions don't change their arguments.

```{r}
x <- c(4, 1, 3)
y <- sort(x)
y
x
```


## Local vs Global {.build}

If `z` is in your global environment and is also defined in the local environment
(inside the function), it inherits the initial global value but does not alter it
in the global env.

```{r}
u <- 3
v <- 7
g <- function(x) {
  x <- x + 2
  u <- u + x
  u
}
```


## Local vs Global {.build}

```{r}
g <- function(x) {
  x <- x + 2
  u <- u + x
  u
}
g(v)
u
v
```


## {.build}

Write a quick function to compute the sum of squares of two vectors of length 1.

```{r}
sum_of_squares <- function(x, y) {
  x^2 + y^2
}
sum_of_squares(2, 5)
```


## Why do we write functions? {.build}

**Eliminate code replication**.

A good function should be:

- short
- perform a single operation
- be well-named

And keep in mind functions can nest inside functions inside functions inside functions
`rep(", inside functions", Inf)`.


## Mini-Project