---
title: 'dplr: Practice'
author: "R Workshop"
output:
  html_document:
    css: ../workshop.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(warning=FALSE, message=FALSE)
```


### Earthquake data

```{r load-quakes, eval = FALSE}
data(quakes)
```

1.   Show a command you could use to find the average (mean) magnitude of all earthquakes that occurred below the median depth of all earthquakes.
1.   How does the average magnitude of these "deep" earthquakes compare to the magnitude of "shallow" (above median depth) earthquakes?
1.   Show the command you would use to get the row(s) with the smallest value in the `depth` column.
1.   Show the command you would use to create a new data frame named `quakes_sorted` that stores the data in sorted order by `depth`.

* * *

### Flights data

```{r load-flights, eval = FALSE}
library(pnwflights14)
data(flights)
```

1. Say I fly very often between PDX airport in Portland and Logan Airport in Boston. Based on the 2013 data, which day of the week would you recommend that I buy a ticket for if I want to minimize delays?

1. What advice do you have for the best airline for this route?


* * *

### Back to Global Warming

Go back to the vector `tData` that you used for the project. Use this data and
maybe `dplyr` to form a dataframe that will make the analysis much more straightforward.
