---
title: "Dataframes: Practice"
author: "R Workshop"
output:
  html_document:
    css: ../workshop.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(warning = FALSE, message = FALSE)
```

For the following exercises, load in the built in earthquake data set.

```{r, eval=FALSE}
data(quakes)
```

1.   How would you find information about this data frame? How many rows and columns does it have?
1.   Based purely on the description of the columns (found with `?quakes`), would you expect a relationship between the 4th and 5th columns? Why or why not?
1.   Show a command you could use to find the average (mean) magnitude of all earthquakes that occurred below the median depth of all earthquakes.
1.   How does the average magnitude of these "deep" earthquakes compare to the magnitude of "shallow" (above median depth) earthquakes?
1.   Show the command you would use to get the row(s) with the smallest value in the `depth` column.



<div id="license">
Adapted from material developed by Olaf Hall-Holt and Paul Roback for CS 125 at St. Olaf College.
</div>