---
title: "Control Flow"
output:
  ioslides_presentation:
    incremental: true
---

## for {.build}

A `for` loop is useful for doing the same task a specific number of times.

```
for (*condition telling us how many times to run the loop*) {
  *task to repeat*
}
```


## for (continued) {.build}

```{r}
total_population <- c()

for (i in 1:nrow(state.x77)) {
  total_population <- c(total_population, state.x77[i, 1])
}
```

We declare an empty object called `total_population`.

Then, we run the `for` loop.  For every row in `state.x77`, we use the combine
that state's population to the `total_population` vector.


## performance {.build}

```{r}
total_population <- c()
system.time(
for (i in 1:nrow(state.x77)) {
  total_population <- c(total_population, state.x77[i, 1])
}
)
```


## performance with microbenchmark {.build}

```{r}
library(microbenchmark)
total_population <- c()
microbenchmark(
for (i in 1:nrow(state.x77)) {
  total_population <- c(total_population, state.x77[i, 1])
}
)
```


## Caveat {.build}

    Premature optimization is the root of all evil.

-Donald Knuth


## for (continued) {.build}

The total population in the United States is...

```{r}
sum(total_population)
```


##  {.build}

Using `microbenchmark` package, compare three flavors of

```{r}
total_population <- c()
for (i in 1:nrow(state.x77)) {
  total_population <- c(total_population, state.x77[i, 1])
}
```

1. **Method I**: As is.
2. **Method II**: Use `append()` instead of `c()`.
3. **Method III**: Initialize `total_population` as a vector of `NA`s of the right length,
then use square bracket indexing to fill them in.

## if {.build}

If statements control the execution of code within them subject to some condition.

The format of an if statement is the following:

```
if (condition) {do these tasks}
```

The if statement is a function which takes one argument: the conditional statement
which is to be evaluated. The bracketed code is then only run when the condition 
has been evaluated as `TRUE`.


## if (continued) {.build}

```{r}
x1 <- runif(1, -1, 1)
x1
```

```{r}
if(x1 > 0){
  print('Positive!')
}
```

In the above example, 'Positive!' is only printed when the conditional evaluates 
to `TRUE`, in this case, when x1 is greater than 0.


## ifelse {.build}

The `ifelse()` command is an extension of the if statement. 

Unlike the if statement, which has one 'argument', the `ifelse()` function has 
three arguments.

The format of an ifelse statement takes the following form:
`ifelse(condition, do if true, do if false)`


## ifelse (continued) {.build}

```{r}
x1 <- runif(1, -1, 1)
x1
```

```{r}
ifelse(x1 > 0, 'Positive', 'Not Positive')
```


##{.build}

Compare run times for two different ways to create a vector that's the same
length of `x` that contains the appropriate strings: `negative` and `positive`.

```{r}
set.seed(32)
x <- runif(1e6, -1, 1)
```

1. **Method I**: Use `ifelse()`.
2. **Method II**: Use square brackets and change-in-place notation.


## while {.build}

`while` loops and executes command until condition is no longer satisfied.

```
while (*condition to be satisfied*) {
*task to be completed for each iteration*
}
```

## while (continued) {.build}

Let's do an example where we want to loop through numbers, starting at 0 and 
ending when we reach 7. We will then add 1 to our number and print it.

```{r}
x <- 0
while (x < 7) {x <- x + 1; print(x);}
```


## next {.build}

`next` stops the processing of the current iteration and advances the index.

It is used within a loop.


##  {.build}

Try an example with a `while` and a `next`. Print numbers starting
at 0 and ending at 7, but don't print the evens. Use next when you hit 
an even number to move to the next iteration.



##  {.build}

Try an example with a `while` and a `next`. Print numbers starting
at 0 and ending at 7, but don't print the evens. Use next when you hit 
an even number to move to the next iteration.

```{r}
x <- 0
while(x < 7) {
  x <- x + 1
  if ( x %% 2 == 0) next
  print(x)
}
```


## break {.build}

`break` breaks out of a loop and control is transferred to the first statement 
that is outside of the current loop.

It can be used in a `for`, `while`, or `repeat` loop.


## break (continued) {.build}

Let's do an example with a `for`. 

```{r}
x <- 0
for (x in 1:7) {
  print(x)
  x <- x + 1
  if (x == 5) break
}
```


## break (continued) {.build}

```{r}
for (x in 1:7) {
  print(x)
  x <- x + 1
}
```


## repeat {.build}

A `repeat` loop contains a single task that loops over and over.  Unlike the `for` 
and `while` loops, it does not have a predefined condition that dictates how many 
times it iterates.

Instead, a `repeat` loops only exits with a `break` statement.


## repeat (continued) {.build}

```
repeat {

  *some task over and over*
  
  if (*condition*) break 
  
}
```

## repeat (continued) {.build}

Generate a random number between 1 and 5.

When we get 2, break.


```{r}
repeat {
  x <- sample(1:5, 1)
  print(x)
  if(x == 2) break
}
```


## warning(s) {.build}

Warnings are functions which indicate there could be trouble.
You can create your own warnings inside your own functions.

```{r eval=FALSE}
myFunction <- function(args...){
  check args...
  warning(condition, message)
  do stuff
}
```


## `warning()` example {.build}

```{r warning = TRUE}
my_function <- function(x, y) {
  if(y == 0) {
    warning('Division by zero')
  }
  return(x/y)
}

my_function(10, 0)
```


## stop {.build}

`stop()` is a way to break out of a loop with an error.

```{r, eval=FALSE}
for (x in 1:20) {     
    x <- x + 1
    if (x %% 6 != 0) {
      print(x)
    } else {     
      stop("x is now divisible by 6")    
    }
}
```

The loop stops and our error message is displayed as soon as `stop()` is called 
(i.e. when x is divisible by 6).


## apply {.build}

A family of functions that applies a function to each element of a vector (or other
structure).

- `apply()`, `sapply()`, `lapply()`, `rapply()`, `mapply()`, `tapply()`, `eapply()`.


```{r, eval = TRUE}
m <- matrix(c(rnorm(30, 0), rnorm(30, 2),
              rnorm(30, 5)), nrow = 30, 
            ncol = 3)
```

**Recall**: A matrix is a vector with a second dimension attribute.  To subset,

```{r, eval = FALSE}
m[rows, columns]
```


## {.build}

Compute the mean of the three columns in the following matrix.

```{r, eval = TRUE}
set.seed(157)
m <- matrix(c(rnorm(30, 0), rnorm(30, 2),
              rnorm(30, 5)), nrow = 30, ncol = 3)
```

- **Method I**: use `apply()`.
- **Method II**: use `colMeans()`.
- **Method III**: use a for loop and the `mean()` function (loop through 3 cols).
- **Method IV**: use nested for loops (loop through all cols and rows).
