---
title: "Vectors"
output:
  ioslides_presentation:
    incremental: true
---

## Before we get started . . . {.build}

The most important function in R?

> **?**, The most important function in R.


## Vectors {.build }

*"Everything is a vector"*

...including scalars, matrices, arrays.


Must be of the same mode:

- integer
- numeric (fp)
- character (string)
- logical (boolean)
- complex

...except for `NA`.


# Generating Vectors

## Generating Vectors {.build}

```{r}
v <- 7
v
c(1:30, v)
l <- c("a", "b", "c", "d")
c(l, v)
```


## : operator {.build}

```{r}
5:8
```

Watch for order of operations.

```{r}
i <- 5
1:i - 1
1:(i - 1)
```


## `seq()` and `rep()` {.build}

```{r}
seq(5, 8)
seq(10, 20, 5)
rep(5, 8)
rep(1:5, 2)
```


# Vector Operations

## Vector Math {.build}

```{r}
x <- c(1, 2, 4)
y <- c(1, 2, 1)
x + y
x + 1
```


## Vector Math {.build}

```{r}
x * y
x %*% y
```


## Vector Indexing {.build}

Use square brackets to pick out sub-vectors.

```{r}
letters[1:4]
letters[c(1, 3)]
z <- c(1, 3, 7)
z[-1]
```


## Vector Indexing {.build}

Consider the vector `x <- c(12,15,8,11,24)`.  How do we create a vector of the
differences between adjacent elements?

```{r}
x <- c(12, 15, 8, 11, 24)
x[-1] - x[-length(x)]
```

Work in vectors instead of loops where possible.


## Vector Subsetting/Filtering {.build}

What's going on here?

```{r}
x <- c(12, 15, 8, 11, 24)
x[x < 10]
x < 10
```

We can index our vector with a logical vector of the same length.


## Subsetting with logicals {.build}

```{r}
x <- c(12, 15, 8, 11, 24)
i <- c(F, F, T, F, F)
x[i]
which(x < 10)
x[x < 10] <- 10
x
```


## Useful fact! {.build}

```{r}
i <- c(F, F, T, F, F)
sum(i)
mean(i)
x <- c(12, 15, 8, 11, 24)
mean(x > 11)
```


## Logical operators {.build}

`<` less than

`<=` less than or equal

`>` greater than

`>=` greater than or equal

`==` exactly equal to

`!=` not equal to

`!x` not equal to x

`x | y` x or y

`x & y` x and y


## {.build}

How can you subset the following vector to exclude both elements less than 10 
and all even elements? (hint: `%%`)

```{r}
x <- c(12, 15, 8, 11, 24)
```

```{r}
x[x > 10 & x %% 2 == 1]
```


# Practice I

## Useful Functions for Vectors

## Vectors Functions {.build}

What is the behavior of the functions below and what would be the output when this
vector is supplied as input?

 \[ \{3.2, 9.16, 4.3, 3.2, 8.9\} \]
 
 - `floor()`
 - `ceiling()`
 - `round()`
 - `abs()`
 - `min()`
 - `max()`


## Vectors Functions {.build}

What is the behavior of the functions below and what would be the output when this
vector is supplied as input?

 \[ \{3.2, 9.16, 4.3, 3.2, 8.9\} \]
 
 - `which.min()`
 - `which.max()`
 - `unique()`
 - `duplicated()`
 - `length()`
 
 
## Sorting {.build}

Consider the vector:

```{r}
x <- c(1,6,5,8)
```

What is the effect of `sort()`?  What about `order()`? 


## {.build}

Given the following output, what is `y`?

```{r, echo = FALSE}
y <- c(-2, 4, 7, -1, 0, 3, 4, -2)
```

```{r}
sort(y)
order(y)
```


## {.build}

```{r}
y <- c(-2, 4, 7, -1, 0, 3, 4, -2)
sort(y)
order(y)
```

- Can you think of a way to use only `order()` to get the same output as 
`sort()`?


# Practice II
