---
title: "Vectors: practice II"
author: "R Workshop"
output:
  html_document:
    css: ../workshop.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(warning=FALSE,message=FALSE)
```


## Sorting Vectors

1.   

In addition to its use on vectors, `order()` can also be used to sort (or
I suppose order!) dataframes.  Consider one of R's preloaded datasets, from
Motor Trend Cars.

```{r}
data(mtcars)
mtcars
```

say you wanted to look through this list of cars and order them in a way in a way
that showcases which ones have the best `mpg`, then, the ones that have the 
highest `hp`.  Using `order()`, rearrange this data frame such that it's ordered
first by `mpg` (low to high) and then by `hp` (high to low).

* * *

## Changing Vector Elements

Please paste the following commands into your Rstudio session, in order to set up the following exercises.

```{r}
require('ggplot2')
load(url('http://www.stolaf.edu/people/olaf/cs125/hw9.RData'))
```

Let's see whether the drawing function is working correctly. Try pasting in the following command, and see what it does:

```{r}
draw(p0x, p0y)
```

If all goes well, a simple sketch of Calvin should appear.

1.   

If you now draw the next set of vectors:

```{r}
draw(p1x, p1y)
```

you will see Calvin and his mirror reflection, except that there is a problem with the data: some garbage points are mixed in with the correct data, resulting in a lot of spikes that reach reach between 75 and 100 in the y direction. Please fix this problem, so that drawing `p1x` with `p1y` results in a clean rendering of Calvin and his reflection.

1.   

Let's take a look at the second example:

```{r}
draw(p2x, p2y)
```

here, we almost have four duplicates of Hobbes in a square configuration, but the upper right drawing is missing. Please duplicate one of the drawings of Hobbes, and place it in the upper right corner (by modifying the `p2x` and `p2y` vectors, that is).

1.   

Let's take a look at the third example:

```{r}
draw(p3x, p3y)
```

here, Calvin is running around a circle, but the drawing closest to the origin is backwards. In particular, the x and y coordinates need to be flipped to restore symmetry. As before, please fix this problem.


<div id="license">
Adapted from material developed by Olaf Hall-Holt and Paul Roback for CS 125 at St. Olaf College.
</div>