---
title: "Vectors: practice I"
author: "R Workshop"
output:
  html_document:
    css: ../workshop.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(warning=FALSE,message=FALSE)
```


## Building vectors

1.   Build the following vectors using the colon operator: `:`.

\[
\{-1, 0, 1, 2\} \\
\{2, 1, 0, -1\}
\]

1.   Build the following vectors using `seq()`.

\[
\{-1, 0, 1, 2\} \\
\{2, 4, 6, 8, 10, 12\} \\
\{2, 3.5, 5, 6.5\}
\]

1.   Build the following vectors using `rep()`.

\[
\{1, 2, 3, 1, 2, 3, 1, 2, 3\} \\
\{1, 1, 2, 2, 3, 3\} \\
\{1, 1, 2, 3, 3\} \\
\{1, 2, 3, 1, 2, 3, 1\} \\
\{1, 1, 2, 2, 3, 3, 1, 1, 2, 2, 3, 3, 1, 1, 2, 2, 3, 3\}
\]

* * *

## Manipulating vectors

Please paste the following commands into your Rstudio session, in order to set 
up the following exercises.

```{r}
require('ggplot2')
load(url('http://www.stolaf.edu/people/olaf/cs125/hw8.RData'))
```


If at some point you want to restore the data, it is fine to paste these commands 
in again whenever you need them. 

Let's see whether the drawing function is working correctly. Try typing in the 
following command, and see what it does:

```{r}
draw(p0x, p0y)
```

If all goes well, a simple sketch of Calvin should appear. If you print `p0x`, 
you will see that it is a vector of numbers, and p0y is another vector of 
numbers, both of the same length. `p0x` and `p0y` contain the x and y coordinates, 
respectively, of the points along the line segments that make up the simple Calvin 
sketch. As an experiment, try the following:

```{r}
draw(p0x, p0y+25)
```

Hopefully, the result is that Calvin appeared to move up a distance of 25 units. 
Before proceeding with the exercises, please think through what is happening with 
these commands, since it will be easier to answer the following questions if you 
clearly understand how the two vectors are related to the plot.

1.   If you now draw the next set of vectors:

```{r}
draw(p1x, p1y)
```

you will see Calvin's buddy Hobbes, except that there is a problem with the data: 
the points that are part of Hobbes on the right side have been mistakenly moved 
up by 25 units. Please fix this problem, so that drawing `p1x` with `p1y` results 
in a more plausible rendering of Hobbes. 

2.   Let's take a look at the second example:

```{r}
draw(p2x, p2y)
```

here, we almost have four duplicates of Calvin in several places, but the upper 
right drawing has been shifted to the left. Please move that part of the drawing 
to the right, so that the four Calvins are symmetrically placed (the same distance 
apart both horizontally and vertically).

3.   Let's take a look at the third example:

```{r}
draw(p3x, p3y)
```

here, we almost have four duplicates of Hobbes in several places, but every other 
point has been moved five units up and to the right, making it almost impossible 
to see Hobbes anywhere. Try to resolve the four duplicates of Hobbes.

<div id="license">
Adapted from material developed by Olaf Hall-Holt and Paul Roback for CS 125 at St. Olaf College.
</div>
